---
title: "PS5"
author: "Simran Bhatti"
date: "2023-02-18"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)

```
# Loading and checking data

```{r, message=FALSE}
gapminder <- read.delim("C:/Users/simra/Documents/Simran Bhatti/INFO 201/ps5-simmi33/gapminder.csv.bz2")

```

The *gapminder* data has `r nrow(gapminder)` rows and `r ncol(gapminder)`


Example of gapminder data:
```{r, message=FALSE}
gapminder %>%
  sample_n(5) %>%
knitr::kable()
```

# Descriptive statistics

```{r}
gapminder %>%
  distinct(gapminder$name) %>%
  count()
```

*There are 250 countries under 'name'*

```{r}
gapminder %>%
  distinct(gapminder$iso2) %>%
  count()
```

*There are 249 countries under 'iso2'*

```{r}
gapminder %>%
distinct(gapminder$iso3) %>%
  count()
```

*There are 253 countries under 'iso3'*

```{r, message=FALSE}
gapminder %>%
  group_by(iso2)%>%
summarize(nameofiso = n_distinct(name), na.rm= TRUE)
```



*There seem to be two iso2 codes that correspond*

```{r}
gapminder %>%
  group_by(name) %>%
  select(iso2, name) %>%
  summarise(num = n_distinct(iso2)) %>%
  arrange(desc(num))

```
*When serching through data, 'NA' countries seem to correspond with the iso-2 codes*

```{r, message=FALSE}
gapminder %>%
  group_by(iso3)%>%
  summarise(nameofiso3 = n_distinct(name), na.rm=TRUE)
```
*There seem to be no iso3 codes that correspond*

```{r}
gapminder %>%
group_by(name) %>%
  select(iso3, name) %>%
  summarise(num =n_distinct(iso3)) %>%
  arrange(desc(num)) 
```
*When searching through data, 'NA' countries seem to correspond with iso-3 codes 4 times. This means that 4 iso-3 code correlate with countries with no name in the data*


```{r, message=FALSE}
min(gapminder$time, na.rm = TRUE)
max(gapminder$time, na.rm = TRUE)
```
The gapminder data set has a maximum of `r max(gapminder$time, na.rm = TRUE)` and a minimum of `r min(gapminder$time, na.rm = TRUE)`

# CO2 Emmisions

```{r}
gapminder %>% 
  filter(is.na(co2), is.na(co2_PC)) %>%
  group_by(time) %>% 
  summarise(missing_co2 = length(co2), missing_co2_pc = length(co2_PC))

```
The years 1960 and 1961 in the data have the most missing data for 'co2' and 'co2_pc' with the total of 60.


```{r, message=FALSE}
gapminder %>% 
group_by(region) %>% 
  summarise(co2_average = mean(co2_PC, na.rm = TRUE))
```